SPARQL on the Open, Decentralised Web

### 2016-08-03 dissertation/introduction.tex
Go through text flow #18

The example opens with prefix declarations, to allow abbreviating
IRIs, e.g. the predicate \rdfterm{foaf:name} is an abbreviation of the
IRI \rdfterm{http://xmlns.com/foaf/0.1/name}. The IRI
\rdfterm{http://www.kjetil.kjernsmo.net/foaf\#me} is a name for this
author. The \rdfterm{a} predicate is standards-defined abbreviation
for \rdfterm{http://www.w3.org/1999/02/22-rdf-syntax-ns\#type}, which
is used for declaring the subject an instance of a class, in this case
\rdfterm{foaf:Person}. Then, the \rdfterm{foaf:made} predicate
is used to link to an author-published version of one of the
publications of this dissertation, and further information is then
provided about this publication, including a link to the co-author. As
we see, semicolons are used to separate statements that share a
subject, commas separate statements with the same subject and
predicate, while periods terminate a statement.

@@

\begin{enumerate}
\item 5 Free Software packages that have been accepted into the Debian
  GNU/Linux operating system.
\item Further 5 packages that already were accepted in Debian have been
  enhanced in the course of this work.
\item Contributions to the code base were accepted by another 5
  external projects.
\end{enumerate}

>>>>

The example opens with prefix declarations, to allow abbreviating
IRIs, e.g. the predicate \rdfterm{foaf:name} is an abbreviation of the
IRI \rdfterm{\nolinkurl{http://xmlns.com/foaf/0.1/name}}. The IRI
\rdfterm{\nolinkurl{http://www.kjetil.kjernsmo.net/foaf\#me}} is a name for this
author. The \rdfterm{a} predicate is standards-defined abbreviation
for \rdfterm{\nolinkurl{http://www.w3.org/1999/02/22-rdf-syntax-ns\#type}}, which
is used for declaring the subject an instance of a class, in this case
\rdfterm{foaf:Person}. Then, the \rdfterm{foaf:made} predicate
is used to link to an author-published version of one of the
publications of this dissertation, and further information is then
provided about this publication, including a link to the co-author. As
we see, semicolons are used to separate statements that share a
subject, commas separate statements with the same subject and
predicate, while periods terminate a statement.

@@

\begin{enumerate}
\item 5 Free Software packages that have been accepted into the recent
  versions of the Debian GNU/Linux operating system.
\item Further 5 packages that already were accepted in Debian have been
  enhanced in the course of this work.
\item Contributions to the code base were accepted by another 5
  external projects.
\end{enumerate}

### 2016-08-03 dissertation/related.tex
Forgot the lampo reference
Go through text flow #18

Interestingly, \cite{papailiou2015graph} opens by noting that many
optimisations used by SQL databases are ineffective in RDF databases,
due to that RDF schema, even if they exist, are not as helpful as the
obligatory schema in SQL databases. This provides an opportunity for
novel research. The solution proposed in \cite{papailiou2015graph}  is not a cache on
a proxy or a mediator like that which is my primary focus. The execution engine
in this case, has access to both the primary indices of the RDF store
and the cache. A key contribution of the paper is a canonical
labelling algorithm that can create a string that will be unique for a
Basic Graph Pattern including optional blocks. This is done by
transforming a SPARQL query to a directed vertex-coloured graph, where
each triple pattern is represented by a vertex, and triple patterns
that share a common variable is linked with an edge. For each vertex,
a label that consists of three IDs are created. Bound terms are
translated to IDs using a dictionary, while variables are translated
to zero. Then, the graph is directed and edges labelled according to
the type of join they represent. The vertex and edge labels are then
translated to non-negative integers (colours) using a sort key, to
form a directed vertex-and-edge-coloured graph that according to the
authors is isomorphic with the original SPARQL. Finally, the graph is
transformed to a simpler directed vertex-coloured graph. With this
graph, the author's method can produce a canonical label for the query
graph.

@@

>>>>

Interestingly, \cite{papailiou2015graph} opens by noting that many
optimisations used by SQL\linebreak[4] databases are ineffective in RDF databases,
due to that RDF schema, even if they exist, are not as helpful as the
obligatory schema in SQL databases. This provides an opportunity for
novel research. The solution proposed in \cite{papailiou2015graph}  is not a cache on
a proxy or a mediator like that which is my primary focus. The execution engine
in this case, has access to both the primary indices of the RDF store
and the cache. A key contribution of the paper is a canonical
labelling algorithm that can create a string that will be unique for a
Basic Graph Pattern including optional blocks. This is done by
transforming a SPARQL query to a directed vertex-coloured graph, where
each triple pattern is represented by a vertex, and triple patterns
that share a common variable is linked with an edge. For each vertex,
a label that consists of three IDs are created. Bound terms are
translated to IDs using a dictionary, while variables are translated
to zero. Then, the graph is directed and edges labelled according to
the type of join they represent. The vertex and edge labels are then
translated to non-negative integers (colours) using a sort key, to
form a directed vertex-and-edge-coloured graph that according to the
authors is isomorphic with the original SPARQL. Finally, the graph is
transformed to a simpler directed vertex-coloured graph. With this
graph, the author's method can produce a canonical label for the query
graph.

@@

Lampo et al. \cite{lampo2011cache} found similar results. They found
that queries with star-shaped groups, that is, triple patterns that
share a common variable in the subject position, take best advantage
of a cache, and suggest rewriting queries to take better advantage of
a cache.
